---
title: "Database Graph Queries"
author: "Jeffrey D Walker, PhD"
date: "7/18/2018"
output: 
  html_document: 
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(jsonlite)
library(ggmap)
library(gridExtra)
library(DBI)

theme_set(theme_bw())

source("load-functions.R")
source("functions.R")

config <- load_config()

con <- dbConnect(
  RPostgreSQL::PostgreSQL(),
  dbname = config$db$dbname,
  host = config$db$host,
  port = config$db$port,
  user = config$db$user
)

barriers <- tbl(con, "barriers") %>%
  select(id, x = x_coord, y = y_coord) %>% 
  collect()
```

# Overview

This document explains how to retrieve a local graph network for one or more barriers from the relational database. Effectively, these queries should provide the same functionality as the `get.graph.tiles()` function.

# Single Target Barrier

This section focuses on retrieving the graph network for a single target, which is selected at random.

```{r}
set.seed(123)
targets <- sample_n(barriers, size = 1)
targets$upgrades <- 0
print(targets)
```

## R Code Algorithm

First, run the R code algorithm using . See `model-algorithm.Rmd` for details about this process.

```{r}
# arguments to graph.linkages()
source = config$tiles$dir       # folder containing tiles
internode = 200                 # internode interval (gives maximum edge length)
tilesize = 400                  # size of tiles
bandwidth = 5000                # bandwidth (m)
search = 1.5                    # search distance (multiple of bandwidth)
cellsize = 30                   # size of cells (m)
fudge = 75                      # how far (m) do we allow x,ys to be from actual culverts?
scaleby = 0.001                 # parameter from CAPS LINKAGES; use 0.001

# get.graph.tiles() arguments
points = as.data.frame(targets) # df of x,y locations of target culverts/dams
nodecosts = NULL                # ??
upgrades = TRUE                 # if TRUE, use points$upgrades for nodecosts in trim.along.graph
```

Compute buffer around target (`maxdist`), which is the farthest distance (m) that kernels can be affected by a given barrier.

```{r}
maxdist <- bandwidth * search + internode + fudge
maxdist
```

Calculate buffer bounding box (`[[xmin, ymin], [xmax, ymax]]`) around the target.

```{r}
xys <- points[, c('x', 'y')]
xys <- cbind(xys, xys - maxdist, xys + maxdist)
colnames(xys) <- c('x', 'y', 'xmin', 'ymin', 'xmax', 'ymax')
xys
```

Given the buffer bouding box, load graph tiles that intersect that box.

```{r}
tiles <- read.table(paste(source, 'graphtiles.txt', sep = ''), sep = '\t', header = TRUE)
q <- outer(tiles$xmin, xys$xmax, '<=') & outer(tiles$ymin, xys$ymax, '<=') & outer(tiles$xmax, xys$xmin, '>=') & outer(tiles$ymax, xys$ymin, '>=')
b <- apply(q, 1, 'any')
tiles <- tiles[b,]
f <- paste(source, sprintf('graph%03d%03d.RDS', tiles$row, tiles$col), sep = '')

nodes <- edges <- matrix(0, 0, 0)

for(i in 1:dim(tiles)[1]) {
  q <- readRDS(f[i])
  nodes <- rbind(nodes, q[[1]])
  edges <- rbind(edges, q[[2]])
}
edges <- unique(edges)
```

The `nodes` matrix contains `r nrow(nodes)` rows.

```{r}
head(nodes)
```

The `edges` matrix contains `r nrow(edges)` rows.

```{r}
head(edges)
```

This map shows the nodes colored by cost, the graph tile boundaries, the location of the target (red circle), and the buffer around the target (red box).

```{r}
nodes %>% 
  ggplot() +
  geom_point(aes(x, y, color = cost)) +
  geom_rect(
    data = tiles,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "black"
  ) +
  geom_point(
    data = targets,
    aes(x, y),
    size = 4, 
    color = "red"
  ) +
  geom_rect(
    data = xys,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "red"
  ) +
  coord_equal() +
  scale_colour_distiller(palette = "Spectral", type = "div") +
  labs(x = NULL, y = NULL)
```

After combining all tiles within the buffer (`maxdist`) of each target, the network is trimmed to a circular area of radius `maxdist` (`r scales::comma(maxdist)` m) using `trim.graph()`, which returns a 2-element list containin `nodes` and `edges`.

```{r}
tiled <- trim.graph(nodes, edges, xys, maxdist, points = TRUE, chatter = FALSE)
str(tiled)
```

There are now `r nrow(tiled$nodes)` nodes and `r nrow(tiled$edges)` edges.

```{r}
tiled$nodes %>% 
  ggplot() +
  geom_point(aes(x, y, color = cost)) +
  geom_rect(
    data = tiles,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "black"
  ) +
  geom_point(
    data = targets,
    aes(x, y),
    size = 4, 
    color = "red"
  ) +
  geom_rect(
    data = xys,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "red"
  ) +
  coord_equal() +
  scale_colour_distiller(palette = "Spectral", type = "div") +
  labs(x = NULL, y = NULL)
```

### Trim Along Graph

Given the nodes and edges within `maxdist` from target (`tiled`), trim the network again based on the resistance kernel using the `trim.along.graph()` function. 

First find the node id corresponding to the target and append to the points data frame.

```{r}
p <- find.nodes(xys, tiled$nodes, fudge)
points <- cbind(points, p)
names(points)[dim(points)[2]] <- 'nodeid'
print(points)
```

Assign node costs to zero for target nodes.

```{r}
nodecosts <- tiled$nodes$cost
nodecosts[match(points$nodeid, tiled$nodes$nodeid)] <- points$upgrades
```

```{r}
trimmed <- trim.along.graph(tiled$nodes, tiled$edges, p, bandwidth = bandwidth, cellsize = cellsize, nodecosts = nodecosts, chatter = FALSE)
```

The number of nodes has now been reduced from `r scales::comma(nrow(tiled$nodes))` to `r scales::comma(nrow(trimmed$nodes))`, and the number of edges reduced from `r scales::comma(nrow(tiled$edges))` to `r scales::comma(nrow(trimmed$edges))`.

```{r}
trimmed$nodes %>% 
  ggplot() +
  geom_segment(
    data = trimmed$edges %>% 
      left_join(
        select(trimmed$nodes, node1 = nodeid, x1 = x, y1 = y),
        by = "node1"
      ) %>% 
      left_join(
        select(trimmed$nodes, node2 = nodeid, x2 = x, y2 = y),
        by = "node2"
      ),
    aes(x = x1, xend = x2, y = y1, yend = y2, color = cost)
  ) +
  geom_point(aes(x, y, color = cost), size = 3) +
  geom_point(
    data = targets,
    aes(x, y),
    size = 4,
    shape = 21,
    color = "red",
    fill = NA
  ) +
  coord_equal() +
  scale_colour_distiller(palette = "Spectral", type = "div") +
  labs(x = NULL, y = NULL)
```

Finally, export to JSON file for testing JavaScript implementation of `trim.along.graph()`.

```{r}
list(
  targets = points,
  input = tiled,
  output = trimmed
) %>% 
  write_json("../test/data/graph-trim-single.json", pretty = TRUE, auto_unbox = TRUE)
```


## Database Algorithm

Using the database, we can perform a simple geospatial query to get all nodes within a distance of a single point.

```{r}
db_nodes <- dbGetQuery(con, "
  SELECT n.id AS nodeid, n.x, n.y, n.cost
  FROM nodes n, barriers b
  WHERE ST_Contains(ST_Buffer(b.geom, $1), n.geom)
  AND b.id=$2
", param = list(maxdist, targets$id))
```

All `db_nodes` (n = `r nrow(db_nodes)`) are in `tiled$nodes` (n = `r nrow(tiled$nodes)`).

```{r}
all(db_nodes$nodeid %in% tiled$nodes$nodeid)
```

But not all `tiled$nodes` are in `db_nodes`.

```{r}
all(tiled$nodes$nodeid %in% db_nodes$nodeid)
```

The differences occur around the edges and are thus not important.

```{r, fig.width=8, fig.height=7}
tiled$nodes %>% 
  ggplot(aes(x, y, color = nodeid %in% db_nodes$nodeid)) +
  geom_point() +
  scale_color_manual(
    "Tile Result In\nDB Result?",
    values = c("TRUE" = "gray50", "FALSE" = "orangered")
  ) +
  coord_equal()
```

Then get edges for the selected nodes.

```{r}
db_edges <- dbGetQuery(con, "
  WITH n AS (
    SELECT n.id AS nodeid
    FROM nodes n, barriers b
    WHERE ST_Contains(ST_Buffer(b.geom, $1), n.geom)
    AND b.id=$2
  )
  SELECT DISTINCT id, start_id, end_id, length, cost, value
  FROM edges, n
  WHERE start_id=n.nodeid OR end_id=n.nodeid
", param = list(maxdist, targets$id)) %>% 
  filter(start_id %in% db_nodes$nodeid, end_id %in% db_nodes$nodeid)
```

```{r}
tile_edgeids <- paste(tiled$edges$node1, tiled$edges$node2, sep = "-")
db_edgeids <- paste(db_edges$start_id, db_edges$end_id, sep = "-")
```

All `db_edges` are in `tiled$edges`

```{r}
all(db_edgeids %in% tile_edgeids)
```

But not vice versa.

```{r}
all(tile_edgeids %in% db_edgeids)
```

Differences are due to the few nodes around the edge. Again, not important.

```{r, fig.width=8, fig.height=7}
db_edges %>%
  left_join(
    db_nodes %>% 
      select(start_id = nodeid, x1 = x, y1 = y),
    by = "start_id"
  ) %>%
  left_join(
    db_nodes %>% 
      select(end_id = nodeid, x2 = x, y2 = y),
    by = "end_id"
  ) %>% 
  ggplot() +
  geom_segment(
    data = tiled$edges %>%
      left_join(
        tiled$nodes %>% 
          select(node1 = nodeid, x1 = x, y1 = y),
        by = "node1"
      ) %>%
      left_join(
        tiled$nodes %>% 
          select(node2 = nodeid, x2 = x, y2 = y),
        by = "node2"
      ),
    aes(x = x1, xend = x2, y = y1, yend = y2),
    color = "orangered"
  ) +
  geom_point(
    data = tiled$nodes,
    aes(x, y),
    size = 1,
    shape = 21,
    color = "orangered",
    fill = NA
  ) +
  geom_segment(
    aes(x = x1, xend = x2, y = y1, yend = y2),
    color = "gray50"
  ) +
  geom_point(
    data = db_nodes,
    aes(x, y),
    size = 1,
    color = "gray50"
  ) +
  coord_equal()
```


# Multiple Target Barriers

This section focuses on retrieving the graph network for multiple targets (n = 4). Three chosen at random, and the fourth selected to be within 500 m of the first target.

```{r}
targets <- barriers %>% 
  filter(id %in% c('c-244895', 'c-244844', 'c-282781', 'c-361794'))
targets$upgrades <- 0
print(targets)
```

## R Code Algorithm

First, run the R code algorithm using . See `model-algorithm.Rmd` for details about this process.

```{r}
# arguments to graph.linkages()
source = config$tiles$dir       # folder containing tiles
internode = 200                 # internode interval (gives maximum edge length)
tilesize = 400                  # size of tiles
bandwidth = 5000                # bandwidth (m)
search = 1.5                    # search distance (multiple of bandwidth)
cellsize = 30                   # size of cells (m)
fudge = 75                      # how far (m) do we allow x,ys to be from actual culverts?
scaleby = 0.001                 # parameter from CAPS LINKAGES; use 0.001

# get.graph.tiles() arguments
points = as.data.frame(targets) # df of x,y locations of target culverts/dams
nodecosts = NULL                # ??
upgrades = TRUE                 # if TRUE, use points$upgrades for nodecosts in trim.along.graph

maxdist <- bandwidth * search + internode + fudge

xys <- points[, c('x', 'y')]
xys <- cbind(xys, xys - maxdist, xys + maxdist)
colnames(xys) <- c('x', 'y', 'xmin', 'ymin', 'xmax', 'ymax')

tiles <- read.table(paste(source, 'graphtiles.txt', sep = ''), sep = '\t', header = TRUE)
q <- outer(tiles$xmin, xys$xmax, '<=') & outer(tiles$ymin, xys$ymax, '<=') & outer(tiles$xmax, xys$xmin, '>=') & outer(tiles$ymax, xys$ymin, '>=')
b <- apply(q, 1, 'any')
tiles <- tiles[b,]
f <- paste(source, sprintf('graph%03d%03d.RDS', tiles$row, tiles$col), sep = '')

nodes <- edges <- matrix(0, 0, 0)

for(i in 1:dim(tiles)[1]) {
  q <- readRDS(f[i])
  nodes <- rbind(nodes, q[[1]])
  edges <- rbind(edges, q[[2]])
}
edges <- unique(edges)
```

The `nodes` matrix contains `r nrow(nodes)` rows.

```{r}
head(nodes)
```

The `edges` matrix contains `r nrow(edges)` rows.

```{r}
head(edges)
```

This map shows the nodes colored by cost, the graph tile boundaries, the location of the target (red circle), and the buffer around the target (red box).

```{r, fig.width=8, fig.height=3}
nodes %>% 
  ggplot() +
  geom_point(aes(x, y, color = cost)) +
  geom_rect(
    data = tiles,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "black"
  ) +
  geom_point(
    data = targets,
    aes(x, y),
    size = 4, 
    color = "red"
  ) +
  geom_rect(
    data = xys,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "red"
  ) +
  coord_equal() +
  scale_colour_distiller(palette = "Spectral", type = "div") +
  labs(x = NULL, y = NULL)
```

After combining all tiles within the buffer (`maxdist`) of each target, the network is trimmed to a circular area of radius `maxdist` (`r scales::comma(maxdist)` m) using `trim.graph()`, which returns a 2-element list containin `nodes` and `edges`.

```{r}
tiled <- trim.graph(nodes, edges, xys, maxdist, points = TRUE, chatter = FALSE)
str(tiled)
```

There are now `r nrow(tiled$nodes)` nodes and `r nrow(tiled$edges)` edges.

```{r, fig.width=8, fig.height=3}
tiled$nodes %>% 
  ggplot() +
  geom_point(aes(x, y, color = cost)) +
  geom_rect(
    data = tiles,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "black"
  ) +
  geom_point(
    data = targets,
    aes(x, y),
    size = 4, 
    color = "red"
  ) +
  geom_rect(
    data = xys,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "red"
  ) +
  coord_equal() +
  scale_colour_distiller(palette = "Spectral", type = "div") +
  labs(x = NULL, y = NULL)
```

Zoomin in on the first two targets that are located close to one another.

```{r, fig.width=8, fig.height=7}
tiled$nodes %>% 
  ggplot() +
  geom_point(aes(x, y, color = cost)) +
  geom_rect(
    data = tiles,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "black"
  ) +
  geom_point(
    data = targets,
    aes(x, y),
    size = 4, 
    color = "red"
  ) +
  geom_rect(
    data = xys,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = NA,
    color = "red"
  ) +
  coord_equal(xlim = c(xys$xmin[[1]], xys$xmax[[1]]), ylim = c(xys$ymin[[1]], xys$ymax[[1]])) +
  scale_colour_distiller(palette = "Spectral", type = "div") +
  labs(x = NULL, y = NULL)
```


### Trim Along Graph

Given the nodes and edges within `maxdist` from target (`tiled`), trim the network again based on the resistance kernel using the `trim.along.graph()` function. 

First find the node id corresponding to the target and append to the points data frame.

```{r}
p <- find.nodes(xys, tiled$nodes, fudge)
points <- cbind(points, p)
names(points)[dim(points)[2]] <- 'nodeid'
print(points)
```

Assign node costs to zero for target nodes.

```{r}
nodecosts <- tiled$nodes$cost
nodecosts[match(points$nodeid, tiled$nodes$nodeid)] <- points$upgrades
```

```{r}
trimmed <- trim.along.graph(tiled$nodes, tiled$edges, p, bandwidth = bandwidth, cellsize = cellsize, nodecosts = nodecosts, chatter = FALSE)
```

The number of nodes has now been reduced from `r scales::comma(nrow(tiled$nodes))` to `r scales::comma(nrow(trimmed$nodes))`, and the number of edges reduced from `r scales::comma(nrow(tiled$edges))` to `r scales::comma(nrow(trimmed$edges))`.

```{r}
trimmed$nodes %>% 
  ggplot() +
  geom_segment(
    data = trimmed$edges %>% 
      left_join(
        select(trimmed$nodes, node1 = nodeid, x1 = x, y1 = y),
        by = "node1"
      ) %>% 
      left_join(
        select(trimmed$nodes, node2 = nodeid, x2 = x, y2 = y),
        by = "node2"
      ),
    aes(x = x1, xend = x2, y = y1, yend = y2, color = cost)
  ) +
  geom_point(aes(x, y, color = cost), size = 3) +
  geom_point(
    data = targets,
    aes(x, y),
    size = 4,
    shape = 21,
    color = "red",
    fill = NA
  ) +
  coord_equal() +
  scale_colour_distiller(palette = "Spectral", type = "div") +
  labs(x = NULL, y = NULL)
```

Finally, export to JSON file for testing JavaScript implementation of `trim.along.graph()`.

```{r}
list(
  targets = points,
  input = tiled,
  output = trimmed
) %>% 
  write_json("../test/data/graph-trim-multiple.json", pretty = TRUE, auto_unbox = TRUE)
```



## Database Algorithm

Using the database, we can perform a simple geospatial query to get all nodes within a distance of a single point.

```{r}
db_nodes <- lapply(targets$id, function (id) {
    dbGetQuery(con, "
      SELECT n.id AS nodeid, n.x, n.y, n.cost
      FROM nodes n, barriers b
      WHERE ST_Contains(ST_Buffer(b.geom, $1), n.geom)
      AND b.id=$2
    ", param = list(maxdist, id))
  }) %>% 
  bind_rows() %>% 
  distinct()
```

All `db_nodes` (n = `r nrow(db_nodes)`) are in `tiled$nodes` (n = `r nrow(tiled$nodes)`).

```{r}
all(db_nodes$nodeid %in% tiled$nodes$nodeid)
```

But not all `tiled$nodes` are in `db_nodes`.

```{r}
all(tiled$nodes$nodeid %in% db_nodes$nodeid)
```

The differences occur around the edges and are thus not important.

```{r, fig.width=8, fig.height=3}
tiled$nodes %>% 
  ggplot(aes(x, y, color = nodeid %in% db_nodes$nodeid)) +
  geom_point() +
  scale_color_manual(
    "Tile Result In\nDB Result?",
    values = c("TRUE" = "gray50", "FALSE" = "orangered")
  ) +
  coord_equal()
```

Zoom into first two targets.

```{r, fig.width=8, fig.height=7}
tiled$nodes %>% 
  ggplot(aes(x, y, color = nodeid %in% db_nodes$nodeid)) +
  geom_point() +
  scale_color_manual(
    "Tile Result In\nDB Result?",
    values = c("TRUE" = "gray50", "FALSE" = "orangered")
  ) +
  coord_equal(xlim = c(xys$xmin[[1]], xys$xmax[[1]]), ylim = c(xys$ymin[[1]], xys$ymax[[1]]))
```

Then get edges for the selected nodes.

```{r}
db_edges <- lapply(targets$id, function (id) {
    dbGetQuery(con, "
      WITH n AS (
        SELECT n.id AS nodeid
        FROM nodes n, barriers b
        WHERE ST_Contains(ST_Buffer(b.geom, $1), n.geom)
        AND b.id=$2
      )
      SELECT DISTINCT id, start_id, end_id, length, cost, value
      FROM edges, n
      WHERE start_id=n.nodeid OR end_id=n.nodeid
    ", param = list(maxdist, id)) %>% 
      filter(start_id %in% db_nodes$nodeid, end_id %in% db_nodes$nodeid)
  }) %>%  
  bind_rows() %>% 
  distinct()
```

```{r}
tile_edgeids <- paste(tiled$edges$node1, tiled$edges$node2, sep = "-")
db_edgeids <- paste(db_edges$start_id, db_edges$end_id, sep = "-")
```

All `db_edges` are in `tiled$edges`

```{r}
all(db_edgeids %in% tile_edgeids)
```

But not vice versa.

```{r}
all(tile_edgeids %in% db_edgeids)
```

Differences are due to the few nodes around the edge. Again, not important.

```{r, fig.width=8, fig.height=3}
db_edges %>%
  left_join(
    db_nodes %>% 
      select(start_id = nodeid, x1 = x, y1 = y),
    by = "start_id"
  ) %>%
  left_join(
    db_nodes %>% 
      select(end_id = nodeid, x2 = x, y2 = y),
    by = "end_id"
  ) %>% 
  ggplot() +
  geom_segment(
    data = tiled$edges %>%
      left_join(
        tiled$nodes %>% 
          select(node1 = nodeid, x1 = x, y1 = y),
        by = "node1"
      ) %>%
      left_join(
        tiled$nodes %>% 
          select(node2 = nodeid, x2 = x, y2 = y),
        by = "node2"
      ),
    aes(x = x1, xend = x2, y = y1, yend = y2),
    color = "orangered"
  ) +
  geom_point(
    data = tiled$nodes,
    aes(x, y),
    size = 1,
    shape = 21,
    color = "orangered",
    fill = NA
  ) +
  geom_segment(
    aes(x = x1, xend = x2, y = y1, yend = y2),
    color = "gray50"
  ) +
  geom_point(
    data = db_nodes,
    aes(x, y),
    size = 1,
    color = "gray50"
  ) +
  coord_equal()
```

Zoom into first two targets.

```{r, fig.width=8, fig.height=7}
db_edges %>%
  left_join(
    db_nodes %>% 
      select(start_id = nodeid, x1 = x, y1 = y),
    by = "start_id"
  ) %>%
  left_join(
    db_nodes %>% 
      select(end_id = nodeid, x2 = x, y2 = y),
    by = "end_id"
  ) %>% 
  ggplot() +
  geom_segment(
    data = tiled$edges %>%
      left_join(
        tiled$nodes %>% 
          select(node1 = nodeid, x1 = x, y1 = y),
        by = "node1"
      ) %>%
      left_join(
        tiled$nodes %>% 
          select(node2 = nodeid, x2 = x, y2 = y),
        by = "node2"
      ),
    aes(x = x1, xend = x2, y = y1, yend = y2),
    color = "orangered"
  ) +
  geom_point(
    data = tiled$nodes,
    aes(x, y),
    size = 1,
    shape = 21,
    color = "orangered",
    fill = NA
  ) +
  geom_segment(
    aes(x = x1, xend = x2, y = y1, yend = y2),
    color = "gray50"
  ) +
  geom_point(
    data = db_nodes,
    aes(x, y),
    size = 1,
    color = "gray50"
  ) +
  coord_equal(xlim = c(xys$xmin[[1]], xys$xmax[[1]]), ylim = c(xys$ymin[[1]], xys$ymax[[1]]))
```
