---
title: "Kernel Algorithm"
author: "Jeffrey D Walker, PhD"
date: "7/17/2018"
output: 
  html_document: 
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(jsonlite)

theme_set(theme_bw())

source("functions.R")
source("load-functions.R")

config <- load_config()

con <- src_postgres(
  dbname = config$db$dbname,
  host = config$db$host,
  port = config$db$port,
  user = config$db$user
)

# load barriers
barriers <- tbl(con, "barriers") %>%
  select(id, x = x_coord, y = y_coord) %>% 
  collect()
```

# Overview

This document describes the algorithm for computing the resistance kernel and delta/effect scores for a single localized graph network.

# Single Target

A single random crossing is first selected from all available crossings.

```{r}
set.seed(1234)
targets <- sample_n(barriers, size = 1)
print(targets)
```

Fetch the localized graph network for this target. 

The initial arguments to `graph.linkages()` are

```{r}
# arguments to graph.linkages
targets = targets               # df of x,y locations of target culverts/dams
source = config$tiles$dir       # folder containing tiles
internode = 200                 # internode interval (gives maximum edge length)
tilesize = 400                  # size of tiles
bandwidth = 5000                # bandwidth (m)
search = 1.5                    # search distance (multiple of bandwidth)
cellsize = 30                   # size of cells (m)
fudge = 75                      # how far (m) do we allow x,ys to be from actual culverts?
scaleby = 0.001                 # parameter from CAPS LINKAGES; use 0.001
multiplier = 1000 * 0.2495 * 24 # multiply scaled kernels to match AQCONNECT
```

First, set the upgrade cost to zero for the target node:

```{r}
targets$upgrades <- 0
```

Then graph network associated with the target crossing using the `get.graph.tiles()` function, which returns a list containing dataframes for the `nodes`, `edges`, and `targets`.

```{r}
d <- get.graph.tiles(points = targets, bandwidth = bandwidth, search = search, tilesize = tilesize, internode = internode, cellsize = cellsize, fudge = fudge, source = source, upgrades = TRUE, chatter = FALSE)
print(d)
```

Now, extract the targets, nodes, and edges data frames.

```{r}
targets <- d$targets    # now have nodeids in targets
nodes <- d$nodes
edges <- d$edges
```

Set the `upgrades` column equal to the node cost, except for the target node (nodeid `r targets$nodeid`) which is set to zero.

```{r}
nodes$upgrades <- nodes$cost
nodes$upgrades[match(targets$nodeid, nodes$nodeid)] <- 0
```

Initialize the base and alternative kernels to zero. Each kernel is a numeric vector with length equal to the number of edges.

```{r}
base <- alt <- rep(0, dim(edges)[1])
```

For each edge, compute the graph kernel given the edge index (`i`), `nodes`, `edges`, and a `multiplier` for scaling the kenerl to match AQCONNECT. The output of `graph.kernel()` is then multiplied by the number of cells within the edge.

```{r, eval=FALSE}
for(i in 1:dim(edges)[1]) {
  base <- base + graph.kernel(i, nodes, edges, multiplier = multiplier)$kern * edges$length[i] / cellsize
  alt <- alt + graph.kernel(i, nodes, edges, nodecost = nodes$upgrades, multiplier = multiplier)$kern * edges$length[i] / cellsize
}
```

The `graph.kernel()` function contains the following code:

```{r, eval=FALSE}
# graph.kernel() arguments
edge <- 1 # equivalent to `i` above
nodecost <- nodes$cost
bench = 62620.5

# initalize kernel to zeros
graph.kern <<- rep(0, dim(edges)[1])

# set value of current edge to account = bandwidth * 3 / cellsize (= 500)
graph.kern[edge] <<- account <- bandwidth * 3 / cellsize

# recursively deplete account in both directions (upstream, downstream)
graph.kernel.spread(edge, 1, account, nodes = nodes, edges = edges, nodecost = nodecost)
graph.kernel.spread(edge, 2, account, nodes = nodes, edges = edges, nodecost = nodecost)

# scale kernel using gaussian scaling as in AQCONNECT/CONNECT_RESCALE
kern <- multiplier * (graph.kern > 0) * dnorm(3 - graph.kern / (bandwidth / cellsize)) / bench

# return scaled and raw kernels
list(kern = kern, rawkern = graph.kern)
```

The `graph.kernel.spread()` function contains the following code

```{r, eval=FALSE}
# arguments
direction = 1 # follow node1
nodecost = nodes$cost
edgecost = edges$cost

# find nodes row corresponding to current edge
n <- match(edges[edge, direction], nodes$nodeid)

# subtract cost of current node from account
a <- account - nodecost[n]

# if account is depleted, done
if (a <= 0)
  return(NULL)

# indices of edges adjacent to node n
e <- (1:dim(edges)[1])[(edges$node1 %in% nodes$nodeid[n]) | (edges$node2 %in% nodes$nodeid[n])]    # adjacent edges

# exclude starting edge
e <- e[e != edge]

# for each adjacent edge
for (i in e) {
  # subtract edge cost from account, limiting to minimum value of zero
  graph.kern[i] <<- max(a - edgecost[i], 0)   # subtract edge cost (but don't go below zero)
  
  # if account depleted, done
  if (graph.kern[i] <= 0)
    return(NULL)

  # spread kernel over adjacent edge i using account of current edge
  graph.kernel.spread(i, (edges$node1[i] == nodes$nodeid[n]) + 1, graph.kern[i], nodes = nodes, edges = edges)
}
```


## First Edge

For the first edge (`i = 1`), initialize kernel with all zeros except first element set to initial account (`bandwidth * 3 / cellsize = 500`).

```{r}
edge <- 1
nodecost <- nodes$cost
bench = 62620.5

# initalize kernel to zeros
graph.kern <- rep(0, dim(edges)[1])

# set value of current edge to account = bandwidth * 3 / cellsize = 500
graph.kern[edge] <- account <- bandwidth * 3 / cellsize

print(graph.kern)
```

The current edge (`i = 1`) is:

```{r}
edges[edge, ]
```

### Spread Upstream

Spread kernel in node1 direction (upstream)

```{r}
# graph.kernel.spread(edge, 1, account, nodes = nodes, edges = edges, nodecost = nodecost)

direction = 1 # follow node1
edgecost = edges$cost

# find nodes row corresponding to current edge
n <- match(edges[edge, direction], nodes$nodeid)

# current node
print(nodes[n, ])
```

Subtract cost of current node from account.

```{r}
a <- account - nodecost[n]
print(a)
```

Account is not depleted (`> 0`) so select edges adjacent to current node.

```{r}
# indices of edges adjacent to node n
e <- (1:dim(edges)[1])[(edges$node1 %in% nodes$nodeid[n]) | (edges$node2 %in% nodes$nodeid[n])]    # adjacent edges
print(e)
```

Exclude starting edge, leaves us with no edges.

```{r}
# exclude starting edge
e <- e[e != edge]
print(e)
```

Kernel thus remains as the initial kernel

```{r}
print(graph.kern)
```

### Spread Downstream

Spread kernel in node2 direction (downstream)

```{r}
# graph.kernel.spread(edge, 2, account, nodes = nodes, edges = edges, nodecost = nodecost)

direction = 2 # follow node2
account   = bandwidth * 3 / cellsize

# find nodes row corresponding to current edge
n <- match(edges[edge, direction], nodes$nodeid)

# current node
print(nodes[n, ])
```

Subtract cost of current node from account.

```{r}
a <- account - nodecost[n]
print(a)
```

Account is not depleted (`> 0`) so select edges adjacent to current node.

```{r}
# indices of edges adjacent to node n
e <- (1:dim(edges)[1])[(edges$node1 %in% nodes$nodeid[n]) | (edges$node2 %in% nodes$nodeid[n])]    # adjacent edges
print(e)
```

Exclude starting edge leaves us with one edge (`row 2`).

```{r}
# exclude starting edge
e <- e[e != edge]
print(e)
print(edges[e, ])
```

Then set kernel value for adjacent edge equal to remaining account (after subtracting nodecost and edgecost).

```{r}
i <- e
print(list(account = account, nodecost = nodecost[n], edgecost = edgecost[i], kern = account - nodecost[n] - edgecost[i]))

graph.kern[i] <- max(a - edgecost[i], 0)
```

Now run `graph.kernel.spread()` for the current (adjacent) edge and go downstream using the kernel for this edge as the starting account.

```{r}
graph.kernel.spread(i, (edges$node1[i] == nodes$nodeid[n]) + 1, graph.kern[i], nodes = nodes, edges = edges)
print(graph.kern)
```

Lastly, add this to the `base` kernel.

```{r}
base <- base + graph.kern
print(base)
```

## Edge Kernels

```{r}
edges_nodes <- edges %>%
  mutate(edgeid = row_number()) %>% 
  left_join(
    nodes %>% 
      select(node1 = nodeid, x1 = x, y1 = y, nodecost1 = cost, nodeupgrades1 = upgrades),
    by = "node1"
  ) %>% 
  left_join(
    nodes %>% 
      select(node2 = nodeid, x2 = x, y2 = y, nodecost2 = cost, nodeupgrades2 = upgrades),
    by = "node2"
  )
```

Here is our network, with nodes and edges colored by baseline cost.

```{r, fig.width=8, fig.height=5}
edges_nodes %>%
  ggplot() +
  geom_segment(aes(x = x1, xend = x2, y = y1, yend = y2, color = cost), size = 1) +
  # geom_label(aes(x = (x1 + x2) / 2, y = (y1 + y2) / 2, label = edgeid)) +
  geom_point(data = nodes, aes(x, y, color = cost), size = 3) +
  geom_text(data = nodes, aes(x, y, label = nodeid), size = 3, hjust = 0, vjust = 0.5, nudge_x = 10) +
  scale_color_distiller(palette = "Spectral") +
  coord_equal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )
```

For each edge in the graph, compute the resistance kernel. Each panel in this figure shows the result of `graph.kernel(i, nodes, edges, multiplier)` for edge `i`.

```{r, fig.width=12, fig.height=8}
edges_kern <- edges_nodes %>% 
  select(id = edgeid) %>% 
  mutate(
    kern = map(id, function (i) {
      k <- graph.kernel(i, nodes, edges, multiplier = multiplier)$kern
      edges_nodes %>% 
        mutate(
          kern = k
        )
    })
  )

edges_kern %>% 
  unnest(kern) %>% 
  ggplot() +
  geom_segment(aes(x = x1, xend = x2, y = y1, yend = y2, color = kern), size = 1) +
  geom_point(data = nodes, aes(x, y), size = 1) +
  scale_color_distiller(palette = "Spectral") +
  facet_wrap(~id) +
  coord_equal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )
```

The overall kernel is then the sum of these individual kernels multiplied by the number of cells within each edge.

```{r}
edges_kern_base <- edges_kern %>% 
  unnest(kern) %>% 
  left_join(
    edges_nodes %>% 
      select(id = edgeid, focal_length = length),
    by = "id"
  ) %>% 
  mutate(
    kern = kern * focal_length / cellsize
  ) %>% 
  group_by(edgeid) %>% 
  summarise(kern = sum(kern)) %>% 
  left_join(edges_nodes, by = c("edgeid")) %>% 
  mutate(kern = kern * length / cellsize)
```

```{r}
r <- graph.linkages(targets, internode = internode, tilesize = tilesize, bandwidth = bandwidth, search = search, cellsize = cellsize, fudge = fudge, source = source, chatter = FALSE, scaleby = scaleby, write = FALSE, multiplier = multiplier)

stopifnot(abs(sum(r$results$kernels$base - edges_kern_base$kern)) < 0.0001)
```

```{r, fig.width=8, fig.height=4}
edges_kern_base %>% 
  ggplot() +
  geom_segment(aes(x = x1, xend = x2, y = y1, yend = y2, color = kern), size = 1) +
  geom_point(data = nodes, aes(x, y), size = 1) +
  scale_color_distiller(palette = "Spectral") +
  coord_equal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )
```

# Export to JSON

Export this expample to JSON for JavaScript app development.

## Targets

```{r, comment=NA}
targets %>% 
  toJSON(pretty = TRUE)
```

## Nodes

```{r, comment=NA}
nodes %>% 
  toJSON(pretty = TRUE)
```

## edges

```{r, comment=NA}
edges %>% 
  toJSON(pretty = TRUE)
```

